On the characteristic and diameter of planar integral point sets \footnote{ The work of the first author was partially supported by the Theoretical Physics and Mathematics Advancement Foundation “BASIS”. }

### 2025-03-12 char-4k-bound/char-4k-bound.tex
[4k3][typo] Just a result of another proofreading round
[4k3] Minor edits after external proofreading

\paragraph{Abstract.}
A point set $M$ in Euclidean plane is called an integral point set in semi-general position if all the distances between the
elements of $M$ are integers, and $M$ does not contain collinear triples.
We improve the lower bound for diameter of such sets in the particular case
when the characteristic of the set is of the form $4k+1$ or $4k+2$.
To achieve that, we combine hyperbolas-based and grid-based toolsets.

\section{Introduction}
	A planar integral point set (IPS) is a set of points in the plane, such that the distance between any pair of its points is an integer, and at least one triple of its points is non-collinear. The latter condition is essential to avoid subsets of a straight line; those are de-facto equivalent to subsets of integer numbers and form a completely different combinatorial object.

	In 1945, Erdős gave an elegant proof~\cite{anning1945integral,erdos1945integral} that every IPS is finite.
	He considered an arbitrary IPS $M$ and chose a non-collinear triple $\{M_1, M_2, M_3\} \in M$ so that any other point $M_0 \in M$ lies on either the straight line $M_1M_2$, the perpendicular bisector to the segment $M_1M_2$, or one of $|M_1M_2|-1$ hyperbolas, where $|M_1M_2|$ stands for length of line segment $M_1M_2$.  Applying the same argument to the line segment $M_1M_3$, Erdös concluded that $\#M \le 4 \cdot |M_1M_2| \cdot |M_1M_3|$, where $\#M$ stands for cardinality of $M$.

	Thus, we can easily infer the lower bound for diameter of an IPS $M$:
	$$
		\operatorname{diam} M \ge \frac{\sqrt{\#M}}{2}.
	$$
	That was the first lower bound for the diameter ever;
	however, Erdős did not only estimate the diameter,
	but also established the toolset for further investigation, the core of which is a system of cofocal hyperbolas.

	Using this toolset, in 2003 Solymosi proved~\cite{solymosi2003note} that
	$$
		\operatorname{diam} M \ge c \cdot \#M.
	$$
	Although Solymosi did not give the constant $c$ explicitly,
	it can be inferred~\cite{our-vmmsh-2018-translit} from his proof that $c=\frac{1}{24}$.

	In~\cite{our-vmmsh-2018-translit} the constant (for $\# M \geq 4$) was improved to $0.3457$
	employing Point Packing in a Square Problem~\cite{markot2005newverified,costa2013valid}.
	In~\cite{my-pps-linear-bound-2019} the approach has been further developed, and the constant has been tightened to $\frac{5}{11}$.
	Finally, in~\cite{my-semi-general-5-4-bound-2019} for IPS $M$ in semi-general position (that is an IPS with no collinear triples) it was proved that
	\begin{equation}
		\label{eq:chebsb_lower_bound}
		\diam M \geq \left(\frac {\# M} 5 \right)^{5/4}
	\end{equation}
	(in assumption that the set has at east 4 points).
	All these lower bounds are based on the Erdős's framework --- cofocal hyperbolas.

	Meanwhile in 1988 Kemnitz introduced~\cite{kemnitz1988punktmengen}
	a characteristic of an IPS, that is a squarefree integer $q$ such as the area
	of a triangle formed by any triple of points from the IPS is comparable with $\sqrt{q}$.
	(Indeed, Kemnitz proved this fact for any point set with rational distances ---
	in contrast with IPS, those can be infinite even if they contain non-collinear triples,
	see~\cite{huff1948diophantine} for an example construction
	and~\cite{solymosi2010question} for some known limitations.)
	In 2000s, Kurz introduced~\cite{kurz2005characteristic} the function $d(2,n)$ that evaluates
	to the minimal possible diameter of planar IPS of cardinality $n$.
	Then Kurz employed Kemnitz's results and found~\cite[Subsection 4.2]{kurz2008minimum} exact values of $d(2,n)$
	up to $n=122$ by exhaustive computer search.
	(Taking the characteristic into consideration allows to boost such search significantly ---
	basically because all triples of an IPS form triangles with equal characteristic.)

@@

	In the present paper, we put the power of Erdős's and Kemnitz's approaches together. Due to squarefree nature of characteristic, it can be of the form $4k+1$, $4k+2$ or $4k+3$, where $k$ is integer.
	We prove that for a particular case of characteristic $4k+1$ and $4k+2$
	the bound~\eqref{eq:chebsb_lower_bound} can be improved to
	\begin{equation}
		\operatorname{diam} M \ge \left( \frac{25}{36} n \right)^{5/4}
		.
	\end{equation}

	In Section 2, we give all the required notions and known results. In Section 3, we discuss some examples of integral point sets in order to demonstrate that none of the classes $4k+1$, $4k+2$, $4k+3$ is too exotic nor pathological. Section 4 is devoted to the connection between Erdős curves and the characteristic. In Section 5, we prove some auxiliary results, and we proceed to the main one in Section 6.

@@

For the sake of brevity, the following notations will be used for sets
of positive integers, non-negative integers and all the integers resp.:
\begin{equation}
	\mathbb{N} = \{1,2,3,4,...\},\quad \mathbb{N}_0 = \mathbb{N} \cup \{0\},
	\quad
	\mathbb{N}_\pm = \{0,\pm 1,\pm 2,\pm 3,\pm 4,...\}
\end{equation}

For a finite set $M$, we will denote its cardinality by $\#M$.

@@

When we say that a set in non-collinear,
we mean that it has at least one non-collinear triple.
If we tighten the condition and require all the triples to be non-collinear,
we get the next definition.

@@

\begin{definition}
	If points $M_1, M_2 \in M \in \mathfrak{M}$,
	then the line segment $M_1 M_2$ is said to be an \emph{edge} of $M$.
\end{definition}

@@

\begin{definition}
	Let $M_1 M_2$ be an edge of $M\in \mathfrak{M}$.
	For an integer $n$, $|n| < |M_1 M_2|$,
	we will call the set of points
	\begin{equation}
		\{M_0 : |M_0 M_1| - |M_0 M_2| = n\}
	\end{equation}
	the \emph{$n$-th Erdős curve}.
\end{definition}

@@

Our attention will be mostly restricted to planar integral point sets
with characteristic of the form $4k+1$ and $4k+2$ in semi-general position.
Thus, if an IPS $M$ satisfies these conditions,
we will write that
$M\in\overline{\mathfrak{M}}'$, and also $M\in\overline{\mathfrak{M}}'_n$ for $n=\#M$.

@@

Let us demonstrate that the classes $4k+1$, $4k+2$ nor $4k+3$
are neither too exotic nor pathological.
In order to do this, we will provide important examples of IPS for each class.

@@

Facher sets with characteristic 1, that are called \emph{semi-crabs},
are investigated in~\cite{antonov2008maximal}.
There are also much more complex IPS with characteristic 1;
for those from $\dot{\mathfrak{M}}_7$, we refer the reader to~\cite{kurz2013constructing}.

@@

\begin{proof}
	Let $|BC| = n$. Then, from the triangle inequality $|AC| < |AB| + |BC|$, we have that $|AC| = m + n - s$, $s \in \mathbb{N}$ and $s \leq m$.
	By the Grid Theorem, we can assume that $A = (-\frac{m}{2}; 0)$, $B = (\frac{m}{2}; 0)$, and $C = (\frac{a}{2m}; \frac{b\sqrt{p}}{2m})$. We find the distance between points $A$ and $C$, and points $B$ and $C$ in coordinates, and form a system of equations
	\begin{equation}
		\label{eq:29}
		\begin{cases}
		\frac{\sqrt{(a+m^2)^2+pb^2}}{2m}&=m+n-s.\\
		\frac{\sqrt{(a-m^2)^2+pb^2}}{2m}&=n
		\end{cases}
	\end{equation}
	Multiply each equation by $2m$ and square both sides of the equations in system~\eqref{eq:29}
	\begin{equation}
		\label{eq:30_pre_weeding_I}
		\begin{cases}
		(a+m^2)^2+pb^2&=4m^2(m+n-s)^2,\\
		(a-m^2)^2+pb^2&=4m^2n^2
		\end{cases}
	\end{equation}
	and then subtract the second equation from the first one:
	\begin{equation}
	(a+m^2)^2-(a-m^2)^2=4m^2(m+n-s)^2-4m^2n^2,
	\label{eq: 31}
	\end{equation}
	$$
	4m^2a=4m^2((m+n-s)^2-n^2),
	$$
	so
	\begin{equation}
		\label{eq:32_pre_weeding_I}
		a=(m-s)(m+2n-s).
	\end{equation}
	Substitute the obtained expression~\eqref{eq:32_pre_weeding_I}
	into the second equation of the system~\eqref{eq:30_pre_weeding_I}:
	\begin{equation}
		((m-s)(m+2n-s)-m^2)^2+pb^2=4m^2n^2,
	\end{equation}
	that immediately gives
	\begin{equation}
		\label{eq:33}
		pb^2=s(2m+2n-s)(2n-s)(2m-s).
	\end{equation}
	Since $s$ is assumed to be odd, let $s=2t+1$, where $t\in\mathbb{N}_0$.
	Then equation \eqref{eq:33} can be rewritten as:
	\begin{equation}
		\label{eq:34.2}
		pb^2=(2t+1)(2m+2n-2t-1)(2n-2t-1)(2m-2t-1)
	\end{equation}
	or
	\begin{eqnarray}
		pb^2+1 &=&4(-4t^4+8t^3m+8t^3n-8t^3-4t^2m^2-12t^2mn+12t^2m- {}\nonumber\\
			 &     &-4t^2n^2+12t^2n-6t^2+4tm^2n-4tmn^2-12tmn+6tm- {}\nonumber\\
			 &     &-4tn^2+6tn-2t+2m^2n-m^2+2mn^2-3mn+m-n^2+n).%{}\nonumber\\
	\label{eq:35}
	\end{eqnarray}
	If equation~\eqref{eq:35} has integer solutions, then both $p$ and $b$ are odd.
	The right-hand side is divisible by 4, which implies that  $pb^2\equiv3\operatorname{mod}{4}$.
	Since $b^{2}\equiv1(\operatorname{mod}4)$, it follows that $p=4k+3$, $k\in \mathbb{N}_{0}$.
\end{proof}

The following theorem follows immediately.

@@

\begin{proof}
	Let $|M_1 M_2| = m$, $m \in \mathbb{N}$, and $n = |M_1 M_3| - |M_2 M_3|$.
	If $m$ is an odd number, then by Lemma~\ref{lem:pre_weeding_I}, the difference $n = |M_1 M_3| - |M_2 M_3|$ equals $m - s$, where $s$ is an even number, is also an odd number.
	Let $m$ be an even number; then $n = |M_1 M_3| - |M_2 M_3| = m - s$ is an even number, as a difference of two even numbers.
\end{proof}

@@

\begin{lemma}
	\label{lem:pre_weeding_II}
	Every set $M = \{A, B, C\} \in \mathfrak{M}_3$ with an even edge $|AB| = 2q$, where $|AC| - |BC| = 2q - s$, $q, s \in \mathbb{N}$, $s$ is an odd number and $s \leq 2q$, has a characteristic of the form $p = 8k + 7$, $k \in \mathbb{N}_0$.
\end{lemma}

@@

\begin{proof}
	Indeed, let $|M_1 M_2|=2q$, $q \in \mathbb{N}$.
	For the $n$-th Erdős curve we have $|M_1 M_3|-|M_2 M_3|=n$.
	According to Lemma~\ref{lem:pre_weeding_II}, the difference $n=|M_1 M_3|-|M_2 M_3|=2q-s$,
	where $s$ is an odd number,
	is also an odd number and $\operatorname{char} M = 8k + 7$, $k\in \mathbb{N}_{0}$.
\end{proof}

@@

	Suppose, to the contrary, that $M={M_1, M_2, M_3, M_4} \in \overline{\mathfrak{M}}'_4$.
	Employ the Grid Theorem and set $M_1(-1;0), M_2(1;0), O(0;0)$.
	Then $M_3$ and $M_4$ lie on the perpendicular bisector to $M_1M_2$.
	Since the distance $|M_3M_4|$ is an integer, the area of triangle $M_1M_3M_4$ is a rational number.
	By Definition~\ref{def:char}, this means that $\operatorname{char}{M}={1}$.
	Then by the Grid Theorem we have $M_3(0;t/4)$. Without loss of generality, let us assume that $t\in\mathbb{N}$. Let $|M_1M_3|=s$. Applying the Pythagorean theorem to triangle $OM_1M_3$, we get
	$$
	1+\frac{t^2}{16}=s^2,
	$$
	or equivalently,
	\begin{equation}
	\label{eq: 42}
	16s^2-t^2=16.
	\end{equation}
	Now, we find the solutions to equation~\eqref{eq: 42} in positive integers.
	To do this, we introduce the substitution: $t=4k$, $k\in\mathbb{N}$.
	Then equation~\eqref{eq: 42} takes the form:
	\begin{equation}
	\label{eq: 43}
	s^2-k^2=1.
	\end{equation}
	Break down the left-hand side of equation~\eqref{eq: 43} into factors:
	\begin{equation}
	\label{eq: 44}
	(s-k)(s+k)=1.
	\end{equation}
	Since $s,k\in\mathbb{N}$, the number 1 can be represented as the product of two integers in two ways: $1\cdot1$ and $-1\cdot(-1)$. In the first case, $(s-k)=1$ and $(s+k)=1$, which gives $s=1,k=0$, which contradicts that $k\in\mathbb{N}$.
	In the second case, $(s-k)=-1$ and $(s+k)=-1$, which gives $s=-1$, $k=0$, which also contradicts that $s,k\in\mathbb{N}$.

@@

\begin{proof}
	Indeed, consider the triangle $ABC$ and let $|BC|=a$, $|AC|=b$.
	Using the triangle inequality $|AB| < |BC|+|AC|$,
	we can represent the length of side $AB$ as $|AB|=a+b-s$, $s \in \mathbb{N}$ and $s \leq a$.
	For $s=1$, the conditions of Lemma~\ref{lem:pre_weeding_I} are satisfied: $s$ is an odd number and $s \leq a$.
	Therefore, the triangle with sides $a$, $b$, and $c=a+b-1$
	has a characteristic of the form $p=4k+3$, $k\in \mathbb{N}_0$.
\end{proof}

@@

\begin{proof}
	Indeed, for $a\geq 3$ the inequality
	$$
		(2a-1)^{1/2} \geq \frac{\sqrt{5}}{\sqrt{3}} a^{1/2}
	$$
	holds, and
	the difference between the left and right-hand sides increases with increasing $a$.
\end{proof}

@@

The following lemma is a classical Erdős-style intersection-enumeration one,
empowered by our Weeding Theorem I.

@@

\begin{proof}
	We will distinguish three cases: two even edges, two odd edges and two edges with different parity.

	Let's consider the first case. Suppose both edges $M_1M_2$ and $M_3M_4$ have even lengths. Then, for each point $N\in M$, one of the following conditions is satisfied:

@@

	Let's consider the second case. Suppose both edges $M_1M_2$ and $M_3M_4$ have odd lengths. Then, for each point $N\in M$, one of the following conditions is satisfied:

@@

	Let's consider the third case. Suppose the lengths of edges $M_1M_2$ and $M_3M_4$ are different. Without loss of generality suppose the edge $M_1M_2$ has even length while edge $M_3 M_4$ has odd length.
	Then, for each point $N\in M$, one of the following conditions is satisfied:

@@

Now we are finally ready to proove our main result.
It improves the bound~\eqref{eq:chebsb_lower_bound}
in our special case.

@@

	If $m \leq \frac{6}{5}p^{2/5}$, then by Lemma~\ref{lem:intersection_enumeration},
	\begin{equation*}
		n \leq |M_1 M_2| \cdot |M_3 M_4| - 2 \leq \frac{36}{25}p^{4/5} - 2
	\end{equation*}
	or equivalently
	\begin{equation}
		p \geq \left(\frac{25(n+2)}{36}\right)^{5/4} \geq \left(\frac{25n}{36}\right)^{5/4}
	\end{equation}
	which is exactly the claim of the theorem.

@@

	According to the results~\eqref{d}, for $3 \leq n \leq 36$ the theorem is true.
	Moreover, it is known that for all $37 \leq n \leq 7^4$, we have $\overline{d}(2, n) > 20\ 000$  and our estimate on the diameter is also true.
	Indeed, $\overline{d}(2, 7^4) \geq \left(\frac{25}{36}\cdot 7^4\right)^{5/4} \approx 10 655$.
	This estimate is weaker than the available numerical results. Therefore, from now on, we can assume that $n \geq 7^4$.

@@

	Another research area that is able to utilize our results fruitfully is the maximization of IPS.
	The Weeding Theorems I and II can sometimes make the exhaustive search up to 4 times faster.

>>>>

\paragraph{Abstract.}
A point set $M$ in Euclidean plane is called an integral point set in semi-general position if all the distances between the
elements of $M$ are integers, and $M$ does not contain collinear triples.
In this paper we improve the lower bound for diameter of such sets in the particular case
when the characteristic of the set is of the form $4k+1$ or $4k+2$.
To achieve that, we combine hyperbolas-based and grid-based toolsets.

\section{Introduction}
	A planar integral point set (IPS) is a set of points in the plane, such that the distance between any pair of its points is an integer, and at least one triple of its points is non-collinear. The latter condition is essential to avoid subsets of a straight line; those are de-facto equivalent to subsets of integers and form a completely different combinatorial object.

	In 1945, Erdős gave an elegant proof that every IPS is finite~\cite{anning1945integral,erdos1945integral} .
	He considered an arbitrary IPS $M$ and chose a non-collinear triple $\{M_1, M_2, M_3\} \in M$ so that any other point $M_0 \in M$ lies on either the straight line $M_1M_2$, the perpendicular bisector to the segment $M_1M_2$, or one of $|M_1M_2|-1$ hyperbolas, where $|M_1M_2|$ stands for length of line segment $M_1M_2$.  Applying the same argument to the line segment $M_1M_3$, Erdös concluded that $\#M \le 4 \cdot |M_1M_2| \cdot |M_1M_3|$, where $\#M$ stands for cardinality of $M$.

	Thus, we can easily infer the lower bound for diameter of an IPS $M$:
	$$
		\operatorname{diam} M \ge \frac{\sqrt{\#M}}{2}.
	$$
	That was the first lower bound for the diameter ever;
	however, Erdős did not only estimate the diameter,
	but also established the toolset for further investigation.
	The core of this approach is a system of cofocal hyperbolas.

	Using this toolset, in 2003 Solymosi proved~\cite{solymosi2003note} that
	$$
		\operatorname{diam} M \ge c \cdot \#M.
	$$
	Although Solymosi did not state the value of $c$ explicitly,
	it can be inferred~\cite{our-vmmsh-2018-translit} from his proof that $c=\frac{1}{24}$.

	In~\cite{our-vmmsh-2018-translit} the constant (for $\# M \geq 4$) was improved to $0.3457$
	employing Point Packing in a Square Problem~\cite{markot2005newverified,costa2013valid}.
	In~\cite{my-pps-linear-bound-2019} the approach has been further developed, and the constant has been tightened to $\frac{5}{11}$.
	Finally, in~\cite{my-semi-general-5-4-bound-2019} for IPS $M$ in semi-general position (that is an IPS with no collinear triples) it was proved that
	\begin{equation}
		\label{eq:chebsb_lower_bound}
		\diam M \geq \left(\frac {\# M} 5 \right)^{5/4}
	\end{equation}
	(assuming that the set has at least 4 points).
	All these lower bounds are based on the Erdős's framework --- cofocal hyperbolas.

	Meanwhile in 1988 Kemnitz introduced~\cite{kemnitz1988punktmengen}
	a characteristic of an IPS, that is a squarefree integer $q$ such as the area
	of a triangle formed by any triple of points from the IPS is comparable with $\sqrt{q}$.
	(Indeed, Kemnitz proved this fact for any point set with rational distances ---
	in contrast with IPS, those can be infinite even if they contain non-collinear triples,
	see~\cite{huff1948diophantine} for an example construction
	and~\cite{solymosi2010question} for some known limitations.)
	In 2000s, Kurz introduced~\cite{kurz2005characteristic} the function $d(2,n)$ that evaluates
	the minimal possible diameter of planar IPS of cardinality $n$.
	Then Kurz employed Kemnitz's results and found the exact values of $d(2,n)$
	up to $n=122$ by exhaustive computer search~\cite[Subsection 4.2]{kurz2008minimum}.
	(Taking the characteristic into consideration allows to boost such search significantly ---
	basically because all triples of an IPS form triangles with equal characteristic.)

@@

	In the present paper, we combine the power of Erdős's and Kemnitz's approaches together. Due to squarefree nature of characteristic, it can be of the form $4k+1$, $4k+2$ or $4k+3$, where $k$ is integer.
	We prove that for a particular case of characteristic $4k+1$ and $4k+2$
	the bound~\eqref{eq:chebsb_lower_bound} can be improved to
	\begin{equation}
		\operatorname{diam} M \ge \left( \frac{25}{36} n \right)^{5/4}
		.
	\end{equation}

	In Section 2, we give all the required notions and known results. In Section 3, we discuss some examples of integral point sets in order to demonstrate that none of the classes $4k+1$, $4k+2$, $4k+3$ is too exotic nor pathological. Section 4 describes a connection between Erdős curves and the characteristic. In Section 5, we prove some auxiliary results, and we proceed to the main one in Section 6.

@@

For the sake of brevity, the following notations are used for sets
of positive integers, non-negative integers and all the integers resp.:
\begin{equation}
	\mathbb{N} = \{1,2,3,4,...\},\quad \mathbb{N}_0 = \mathbb{N} \cup \{0\},
	\quad
	\mathbb{N}_\pm = \{0,\pm 1,\pm 2,\pm 3,\pm 4,...\}
\end{equation}

For a finite set $M$, we denote its cardinality by $\#M$.

@@

When we say that a set is non-collinear,
we mean that it has at least one non-collinear triple.
If we tighten the condition and require all the triples to be non-collinear,
we obtain the next definition.

@@

\begin{definition}
	For any two points $M_1, M_2 \in M \in \mathfrak{M}$,
	the line segment $M_1 M_2$ is said to be an \emph{edge} of $M$.
\end{definition}

@@

\begin{definition}
	Let $M_1 M_2$ be an edge of $M\in \mathfrak{M}$.
	For an integer $n$, $|n| < |M_1 M_2|$,
	we call the set of points
	\begin{equation}
		\{M_0 : |M_0 M_1| - |M_0 M_2| = n\}
	\end{equation}
	the \emph{$n$-th Erdős curve}.
\end{definition}

@@

Our attention will be mostly restricted to planar integral point sets
with characteristic of the form $4k+1$ and $4k+2$ in semi-general position.
Thus, if an IPS $M$ satisfies these conditions,
we write that
$M\in\overline{\mathfrak{M}}'$, and also $M\in\overline{\mathfrak{M}}'_n$ for $n=\#M$.

@@

Let us demonstrate that the classes $4k+1$, $4k+2$ nor $4k+3$
are neither too exotic nor pathological.
In order to do this, we provide important examples of IPS for each class.

@@

Facher sets with characteristic 1, that are called \emph{semi-crabs},
are investigated in~\cite{antonov2008maximal}.
There are also significantly more complex IPS with characteristic 1;
for those from $\dot{\mathfrak{M}}_7$, we refer the reader to~\cite{kurz2013constructing}.

@@

\begin{proof}
	Let $|BC| = n$. Then, from the triangle inequality $|AC| < |AB| + |BC|$, we have that $|AC| = m + n - s$, $s \in \mathbb{N}$ and $s \leq m$.
	By the Grid Theorem, we can assume that $A = (-\frac{m}{2}; 0)$, $B = (\frac{m}{2}; 0)$, and $C = (\frac{a}{2m}; \frac{b\sqrt{p}}{2m})$. We find the distance between points $A$ and $C$, and points $B$ and $C$ in coordinates, and form a system of equations
	\begin{equation}
		\label{eq:29}
		\begin{cases}
		\frac{\sqrt{(a+m^2)^2+pb^2}}{2m}&=~m+n-s,\\
		\frac{\sqrt{(a-m^2)^2+pb^2}}{2m}&=~n
		.
		\end{cases}
	\end{equation}
	Multiplying each equation by $2m$ and squaring both sides, we obtain
	\begin{equation}
		\label{eq:30_pre_weeding_I}
		\begin{cases}
		(a+m^2)^2+pb^2&=~4m^2(m+n-s)^2,\\
		(a-m^2)^2+pb^2&=~4m^2n^2
		.
		\end{cases}
	\end{equation}
	Subtract the second equation from the first one:
	\begin{equation}
	(a+m^2)^2-(a-m^2)^2=4m^2(m+n-s)^2-4m^2n^2,
	\label{eq: 31}
	\end{equation}
	$$
	4m^2a=4m^2((m+n-s)^2-n^2),
	$$
	so
	\begin{equation}
		\label{eq:32_pre_weeding_I}
		a=(m-s)(m+2n-s).
	\end{equation}
	Substitute the expression~\eqref{eq:32_pre_weeding_I}
	into the second equation of~\eqref{eq:30_pre_weeding_I}:
	\begin{equation}
		((m-s)(m+2n-s)-m^2)^2+pb^2=4m^2n^2.
	\end{equation}
	This immediately gives
	\begin{equation}
		\label{eq:33}
		pb^2=s(2m+2n-s)(2n-s)(2m-s).
	\end{equation}
	Since $s$ is assumed to be odd, let $s=2t+1$, where $t\in\mathbb{N}_0$.
	Then equation \eqref{eq:33} can be rewritten as:
	\begin{equation}
		\label{eq:34.2}
		pb^2=(2t+1)(2m+2n-2t-1)(2n-2t-1)(2m-2t-1)
	\end{equation}
	or
	\begin{eqnarray}
		pb^2+1 &=&4(-4t^4+8t^3m+8t^3n-8t^3-4t^2m^2-12t^2mn+12t^2m- {}\nonumber\\
			 &     &-4t^2n^2+12t^2n-6t^2+4tm^2n-4tmn^2-12tmn+6tm- {}\nonumber\\
			 &     &-4tn^2+6tn-2t+2m^2n-m^2+2mn^2-3mn+m-n^2+n).%{}\nonumber\\
	\label{eq:35}
	\end{eqnarray}
	If equation~\eqref{eq:35} has integer solutions, then both $p$ and $b$ are odd.
	The right-hand side is divisible by 4, which implies that  $pb^2\equiv3\operatorname{mod}{4}$.
	Since $b^{2}\equiv1(\operatorname{mod}4)$, it follows that $p=4k+3$, $k\in \mathbb{N}_{0}$.
\end{proof}

The next theorem follows immediately.

@@

\begin{proof}
	Let $|M_1 M_2| = m$, $m \in \mathbb{N}$, and $n = |M_1 M_3| - |M_2 M_3|$.
	If $m$ is odd, then by Lemma~\ref{lem:pre_weeding_I}, the difference $n = |M_1 M_3| - |M_2 M_3|$ equals $m - s$, where $s$ is even. Thus, $n$ is odd.
	If $m$ is even, then $n = |M_1 M_3| - |M_2 M_3| = m - s$ is even, as a difference of two even numbers.
\end{proof}

@@

\begin{lemma}
	\label{lem:pre_weeding_II}
	Every set $M = \{A, B, C\} \in \mathfrak{M}_3$ with an even edge $|AB| = 2q$, where $|AC| - |BC| = 2q - s$, $q, s \in \mathbb{N}$, $s$ is odd and $s \leq 2q$, has a characteristic of the form $p = 8k + 7$, $k \in \mathbb{N}_0$.
\end{lemma}

@@

\begin{proof}
	Indeed, let $|M_1 M_2|=2q$, $q \in \mathbb{N}$.
	For the $n$-th Erdős curve we have $|M_1 M_3|-|M_2 M_3|=n$.
	According to Lemma~\ref{lem:pre_weeding_II}, the difference $n=|M_1 M_3|-|M_2 M_3|=2q-s$,
	where $s$ is odd,
	is also odd and $\operatorname{char} M = 8k + 7$, $k\in \mathbb{N}_{0}$.
\end{proof}

@@

	Suppose, to the contrary, that $M={M_1, M_2, M_3, M_4} \in \overline{\mathfrak{M}}'_4$.
	We employ the Grid Theorem and set $M_1(-1;0), M_2(1;0), O(0;0)$.
	Then $M_3$ and $M_4$ lie on the perpendicular bisector to $M_1M_2$.
	Since the distance $|M_3M_4|$ is an integer, the area of triangle $M_1M_3M_4$ is a rational number.
	By Definition~\ref{def:char}, this means that $\operatorname{char}{M}={1}$.
	Then by the Grid Theorem we have $M_3(0;t/4)$. Without loss of generality, let us assume that $t\in\mathbb{N}$. Let $|M_1M_3|=s$. Applying the Pythagorean theorem to triangle $OM_1M_3$, we obtain
	$$
	1+\frac{t^2}{16}=s^2,
	$$
	or equivalently,
	\begin{equation}
	\label{eq: 42}
	16s^2-t^2=16.
	\end{equation}
	To find the solutions to equation~\eqref{eq: 42} in positive integers,
	we introduce the substitution: $t=4k$, $k\in\mathbb{N}$.
	Then equation~\eqref{eq: 42} takes the form:
	\begin{equation}
	\label{eq: 43}
	s^2-k^2=1.
	\end{equation}
	Factorising the left-hand side of equation~\eqref{eq: 43} yields:
	\begin{equation}
	\label{eq: 44}
	(s-k)(s+k)=1.
	\end{equation}
	Since $s,k\in\mathbb{N}$, the number 1 can be represented as the product of two integers in two ways: $1\cdot1$ and $-1\cdot(-1)$. In the first case, $(s-k)=1$ and $(s+k)=1$, which gives $s=1,k=0$. This contradicts that $k\in\mathbb{N}$.
	In the second case, $(s-k)=-1$ and $(s+k)=-1$, which gives $s=-1$, $k=0$. This also contradicts that $s,k\in\mathbb{N}$.

@@

\begin{proof}
	Consider the triangle $ABC$ and let $|BC|=a$, $|AC|=b$.
	Using the triangle inequality $|AB| < |BC|+|AC|$,
	we can represent the length of side $AB$ as $|AB|=a+b-s$, $s \in \mathbb{N}$ and $s \leq a$.
	For $s=1$, the conditions of Lemma~\ref{lem:pre_weeding_I} are satisfied: $s$ is odd and $s \leq a$.
	Therefore, the triangle with sides $a$, $b$, and $c=a+b-1$
	has a characteristic of the form $p=4k+3$, $k\in \mathbb{N}_0$.
\end{proof}

@@

\begin{proof}
	For $a\geq 3$ the inequality
	$$
		(2a-1)^{1/2} \geq \frac{\sqrt{5}}{\sqrt{3}} a^{1/2}
	$$
	holds, and
	the difference between the left and right-hand sides increases with increasing $a$.
\end{proof}

@@

The following result is a classical Erdős-style intersection-enumeration lemma,
empowered by our Weeding Theorem I.

@@

\begin{proof}
	Consider three cases: two even edges, two odd edges and two edges with different parity.

	\paragraph{Case 1.} Suppose that both edges $M_1M_2$ and $M_3M_4$ have even lengths. Then, for each point $N\in M$, one of the following conditions is satisfied:

@@

	\paragraph{Case 2.} Suppose that both edges $M_1M_2$ and $M_3M_4$ have odd lengths. Then, for each point $N\in M$, one of the following conditions is satisfied:

@@

	\paragraph{Case 3.} Suppose that the lengths of edges $M_1M_2$ and $M_3M_4$ are different. Without loss of generality suppose the edge $M_1M_2$ has even length while edge $M_3 M_4$ has odd length.
	Then, for each point $N\in M$, one of the following conditions is satisfied:

@@

Now we are finally ready to prove our main result.
It improves the bound~\eqref{eq:chebsb_lower_bound}
in our special case.

@@

	If $m \leq \frac{6}{5}p^{2/5}$, then by Lemma~\ref{lem:intersection_enumeration},
	\begin{equation*}
		n \leq |M_1 M_2| \cdot |M_3 M_4| - 2 \leq \frac{36}{25}p^{4/5} - 2
	\end{equation*}
	or equivalently
	\begin{equation}
		p \geq \left(\frac{25(n+2)}{36}\right)^{5/4} \geq \left(\frac{25n}{36}\right)^{5/4}
	\end{equation}
	which is exactly the assertion of the theorem.

@@

	According to the results~\eqref{d}, for $3 \leq n \leq 36$ the assertion follows.
	Moreover, it is known that for all $37 \leq n \leq 7^4$, we have $\overline{d}(2, n) > 20\ 000$  and our estimate on the diameter also holds.
	Indeed, $\overline{d}(2, 7^4) \geq \left(\frac{25}{36}\cdot 7^4\right)^{5/4} \approx 10 655$.
	This estimate is weaker than the available numerical results. Therefore, from now on, we can assume that $n \geq 7^4$.

@@

	Another research area that is able to utilize our results fruitfully is the maximization of IPS.
	The Weeding Theorems I and II can sometimes make the exhaustive search up to four times faster.
